### 2016-06-17 docs/documentation/documentation.tex
hopefully endversion documentation

\documentclass[11pt, a4paper]{article}

\usepackage{amssymb, amsmath}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{gensymb}

\author{Manuel Haid \and Thomas Mauerhofer \and Matthias Wölbitsch}
\title{\texttt{timeline} Documentation}

% content:
% it's about „Technical and code“ documentation (code)
% Requirements
% Libraries / frameworks used
% Which step/code does what in detail
% preferably with flow charts
% One should be able to extend or adapt your code easily after reading your documentation

\begin{document}

\maketitle
\tableofcontents

\section{Introduction} \label{sec:intoduction}

In this section we want to explain how to start our project.

\begin{itemize}
\item
\textbf{go to the vagrant folder:} \texttt{cd vagrant/}
\item
\textbf{start vagrant:} \texttt{vagrant up} (the first time can take a long time, because of building the source)
\item
\textbf{start our program:} after vagrant has been started, you can access it from the outside with port 5000. 
Start a browser and go to \texttt{http://localhost:5000/}
\end{itemize}

Now our example application should start and you can explore it.

\section{Requirements} \label{sec:requirments}

\begin{itemize}
 \item \textgreater= Python 3.5
\end{itemize}

\section{Dependencies}\label{sec:dependencies}

This section contains a description of all used external dependencies (e.g. libraries) and why they were used in this project.
The python dependencies can be installed using \texttt{pip}, the Python package manager, using the \texttt{requirements.txt} file. 
It is recommended to use a virtual environment when developing for this project to avoid conflicts with other system-wide installed versions of the same dependencies.

\subsection{Back-end (Python)}

\begin{description}
 \item[Flask] 
 Flask is a microframework (simple but extendable) for web development. 
 It allows RESTful request dispatching and is well documented. 
 In this project Flask is used for the rendering of the front-end using its template engine and providing an API for the supported visualizations.

 \item[Pandas] 
 Pandas is a widely used data analysis library. 
 It provides data structures to efficiently store and query in-memory data. 
 All temporal data (i.e. all time series) in this project is stored in Pandas DataFrame or Series objects.

 \item[NumPy] 
 NumPy is a powerful package for scientific computing. 
 It provides efficient data structures and algorithms to operate on them. 
 However, in this project this library is only used to sample random data for the example data sets and is therefore not very important. 
\end{description}

\subsection{Front-end (HTML5)}

\begin{description}
 \item[jQuery] jQuery is the feature rich JavaScript library to simplify client-side scripting. 
 In this project it is mainly used for AJAX calls.
 \item[MetricsGraphics.js] MetricsGraphics.js is used to render the time series visualizations.
 It is based on the well-known D3.js graphic library and provides all necessary primitives and features to plot the required figures.
 \item[Bootstrap] This CSS framework is used to for its beautiful design templates to provide a familiar and responsive user interface.
 Furthermore, some bootstrap add-ons were used for additional widgets (e.g. the bootstrap datetime picker).
\end{description}

\section{Modules}\label{sec:modules}

\subsection{APP}
This module simply creates the Flask object and resisters the routes for the front-end and for the API.
The Flask object must be used to start timeline using the \texttt{run} method of the app object. 
The source code is located in the \texttt{app.py} file.

\subsection{API}
The task of this module is to take requests from the front-end in form of AJAX calls and return the visualization primitives for the MetricsGraphics.js library as JSON objects.
The source code is located in the \texttt{api.py} file.

\subsubsection*{Time Series Plot: \texttt{api/time\_plot/\{time\_series\_id\}}}
\begin{itemize}
 \item[] \textbf{Method:} GET 
 \item[] \textbf{Description:} This route create the visualization primitive for a simple time plot (i.e. a time plot of only one time series).
 It takes the identifier of the time series as part of the URL (e.g. \texttt{api/time\_plot/42}) and can take additional parameter.
 It may return a HTTP status code 404 if no time series with the given identifier exists and a HTTP status code 400 for invalid values for the optional GET parameter.
 \item[] \textbf{Additional Parameter:} All of the following parameter are optional.
 It is possible to only retrieve a slice of the time series using the \texttt{start\_date} and/or \texttt{end\_date} parameter.
 Each of this parameter must be an integer, which represents to UNIX time stamp of the start time of the slice and the end time of the slice, respectively. 
 Additional it is possible to add the graphs of the rolling mean and/or the rolling standard deviation to the plot.
 The \texttt{rolling\_mean\_window} and \texttt{rolling\_std\_window} parameter must be integer that represent the number of samples that is used to calculate the running mean and standard deviation, respectively.
\end{itemize}

\subsubsection*{Time Series Plots: \texttt{api/time\_plots}}
\begin{itemize}
 \item[] \textbf{Method:} GET 
 \item[] \textbf{Parameter:} list of time series ids
 \item[] \textbf{Description:} This route create the visualization primitive for a simple time plot (i.e. a time plot of all time series).
 It may return a HTTP status code 404 if one time series data doesn't exist.
\end{itemize}

\subsubsection*{Live Time Series Plot: \texttt{api/live\_plot/\{live\_time\_series\_id\}}}
\begin{itemize}
 \item[] \textbf{Method:} GET 
 \item[] \textbf{Parameter:} last received
 \item[] \textbf{Description:} This route create the visualization primitive for a simple live time plot.
 The first request from this function is without an parameter, because it only gets the data. These function with the optional 
 parameter \texttt{last\_received} signaled only an update from the data.
 It may return a HTTP status code 404 if no time series with the given identifier exists.
\end{itemize}

\subsubsection*{ACF Plot: \texttt{api/acf\_plot/\{time\_series\_id\}}}
\begin{itemize}
 \item[] \textbf{Method:} GET 
 \item[] \textbf{Parameter:} max lag, scale
 \item[] \textbf{Description:} This route create the visualization primitive for a simple acf plot.
 It may return a HTTP status code 404 if no time series with the given identifier exists and a HTTP status code 400 for invalid values for the optional GET parameter.
 It is possible to only retrieve a slice of the time series using the \texttt{max\_lag} parameter. The parameter must be an integer.
 The second parameter \texttt{scale} is optional, if it is null, the plot will be autoscaled.
\end{itemize}

\subsubsection*{Forecasting Plot: \texttt{api/forecasting\_plot/\{forecast\_id\}}}
\begin{itemize}
 \item[] \textbf{Method:} GET 
 \item[] \textbf{Parameter:} none
 \item[] \textbf{Description:} This route create the visualization primitive for a simple forecasting plot.
 It may return a HTTP status code 404 if no time series with the given identifier exists.
\end{itemize}

\subsection{DATA SETS}
The task of this module is, to define the class that describes the data sets for the time series, live data and forecast data.
It also has functions to add and get the data from the right classes like described below.
The source code is located in the \texttt{data\_sets.py} file.

\subsubsection*{Class TimeSeries:}
\begin{itemize}
 \item[] \textbf{Description:} The member variables from the class TimeSeries are the autogenerated id, a name, a description, the data and optional the legend. 
 It also has functions to get several informations like the number of samples, the start date, the end date and the period.
\end{itemize}

\subsubsection*{Class LiveTimeSeries:}
\begin{itemize}
 \item[] \textbf{Description:} These class is derived from the class TimeSeries. It only has additional functions to get and to update the data.
\end{itemize}

\subsubsection*{Class TimeSeriesForecast:}
\begin{itemize}
 \item[] \textbf{Description:} These class is also derived from the class TimeSeries. It only has additional member variables like 
 the training data, the forecasted data, the test data and the validation split.
\end{itemize}

\subsubsection*{Register Data Sets:}
\begin{itemize}
 \item[] \textbf{Description:} In this section, we have three functions to add the data sets to the \texttt{time\_series}, \texttt{live\_time\_series} or \texttt{forecast} dictionaries.
\end{itemize}

\subsubsection*{Get Data Sets:}
\begin{itemize}
 \item[] \textbf{Description:} In this section, we have different functions to get the data sets (\texttt{time\_series}, \texttt{live\_time\_series} or \texttt{forecast}). 
 There are functions to get one data set (with an id) or you can get all data sets.
\end{itemize}

\subsection{FRONT END}
The task of this module is to render the different html templates, for example the info.html, index.html, multi\_time\_plots.html, live\_info.html, etc.

\subsection{VISUALIZATION}
The task of this module is to build the data so that the visualization library in the front end can show it.
The source code is located in the \texttt{visualizations.py} file.

\subsubsection*{Time Series Plot:}
\begin{itemize}
 \item[] \textbf{Description:} This function generate the data to visualize a simple time series plot, with the optional parameter start and end date.
An optional feature is, that if the data set is a list, multiple data sets will be used, otherwise only one data set will be used.
Afterwards the title will be set and we return a object with a title, the x accessor, the date, the y accessor, the value,
the data and the legend.
\end{itemize}

\subsubsection*{Add Rolling Mean:}
\begin{itemize}
 \item[] \textbf{Description:} This function adds a rolling mean line to the existing plot. 
 We calculate the rolling mean and add the data object with the data and the legend to the existing data set.
\end{itemize}

\subsubsection*{Add Rolling Std:}
\begin{itemize}
 \item[] \textbf{Description:} This function adds a rolling standard deviation to the existing plot. 
 We calculate the rolling standard deviation and add the data object with the data and the legend to the existing data set.
\end{itemize}

\subsubsection*{Auto Correlation Plot:}
\begin{itemize}
 \item[] \textbf{Description:} This function generates an auto correlation for an existing data set. The \texttt{max\_lag} parameter 
 describes the max number of correlation steps.
\end{itemize}

\subsubsection*{Forecasting Eval Plot:}
\begin{itemize}
 \item[] \textbf{Description:} This function generates the visualisation primitives.
\end{itemize}

\subsubsection*{Build Data Object:}
\begin{itemize}
 \item[] \textbf{Description:} This function converts the time series data to the right format, so that the visualization library can use it. The object consists of a date and a value.
\end{itemize}

\end{document}